### 2021-03-12 manuscript.Rmd
dca - no harm

---
title: "Predicting prolonged apneic episodes using machine learning during nurse-administered procedural sedation"
bibliography: "Extras/references.bib"
---

## Introduction

A capnography waveform displays the level of expired carbon dioxide (CO~2~) over time to show changes in concentrations throughout the respiratory cycle. Capnography waveform abnormalities assist in the detection and diagnosis of specific conditions, such as partial airway obstruction and apnea. For this reason, implementing capnography into practice for respiratory monitoring is considered a high priority to improve patient safety by leading authorities including peak professional organisations for anaesthesia in Canada, the United States and Europe.[@apfelbaum2018; @hinkelbeinEuropeanSocietyAnaesthesiology2018; @Dobson2018]

A potential consequence to consider in regard to the implementation of capnography for patient monitoring in inpatient wards is alarm fatigue. Alarm fatigue has been linked to patient deaths resulting from clinicians becoming desensitized to alarms leading to delayed responses to clinical deterioration.[@chopra2014redesigning] Deciphering which capnography waveform abnormalities deserve intervention (and therefore alarms to signal the event to clinicians) from those that do not is an essential step towards successful implementation of this technology into practice. For example, triggering alarms after short periods of apnea leads to frequent interruptions and potentially increases risk of alarm fatigue. Conversely, only intervening once an apneic period reaches a longer threshold negates the potential benefits of capnography on patient safety by improving ventilation. In practice, two alternative strategies for capnography alarm management are typically used. The 'aggressive' strategy is for alarms to be triggered after short periods of apnea (15 seconds). The 'conservative' approach triggers the alarm only once the patient has been apneic for a prolonged period (30 seconds).

It is possible that capnography monitor alarm management may be improved upon by using machine learning to create a ‘smart alarm’ that can alert clinicians for apneic events that are predicted to be prolonged. Such an approach aligns with a call from The Society for Critical Care Medicine Alarm and Alert Fatigue Task Force, that machine learning techniques should be used to advance the quality of alerts that clinicians receive and to individualize alert delivery based on clinician response characteristics, such as alert frequency and severity.

For the aggressive strategy, only triggering an alarm for apneas predicted to be prolonged would reduce the total alarm burden and potentially reduce the risk of alarm fatigue. The downside of applying a machine learning approaching to the aggressive apnea treatment strategy would be that some prolonged apneas may not receive early intervention if the model incorrectly predicted that the apnea will not last for >30 seconds (i.e. the false negatives). For the conservative strategy, triggering an alarm at the 15-second timepoint for apneas predicted to be prolonged could reduce the total time of apnea, as treatment could be initiated earlier. The downside of applying a machine learning approaching to the conservative apnea treatment strategy would be that there is the potential that it could increase the total alarm burden if the ratio of false positives (predicting the apnea will last for >30 seconds but it does not) to true positives (correctly predting at the 15 second timepoint that the apnea will last for >30 seconds) is high. The aim of this study was to determine the accuracy of machine learning models for predicting at the 15 second timepoint if a period of apnea will persist for 30 seconds or more with the goal of determining whether operationalizing these predictions in practice as alarm triggers would be beneficial.

# Methods

A secondary analysis of a prospective observational study was undertaken. Results for the primary aim of the observational study are reported elsewhere.[@conway2019sequence] All participants provided written informed consent and the study was approved by human research ethics committees (UCH HREC 1614; SVHAC HREC 16/26; QUT 1600000641). The study was undertaken in accordance with the Australian National Statement on Ethical Conduct in Human Research[@national2007] and it was registered prospectively (ACTRN12616001132437).

### Prediction Goal

The prediction goal was to classify apneic events at the 15-second time-point as being either short (i.e. terminated prior to 30 seconds) or prolonged (persisted for 30 seconds or longer). The prediction algorithm was compared against a default alarm setting for capnography monitors, which triggers an alarm at 30 seconds of apnea.

### Participants

Participants in the study were consecutive adult patients scheduled to undergo an elective procedure in the cardiac catheterization laboratory with moderate sedation. Patients with severe cognitive impairment (due to inability to provide informed consent) or inability to understand and speak English (if an interpreter was unavailable) were excluded.

### Sedation and monitoring

The sedation regimen used for patients included in this study comprised bolus doses of intravenous midazolam and fentanyl. Sedation was administered by nurses who were trained in advanced life support. Routine clinical monitoring included cardiac rhythm and pulse oximetry monitoring as well as non-invasive blood pressure measurements every 5-10 minutes. The Respironics LoFlo sidestream CO~2~ sensor was used for capnography monitoring. A CO~2~ sampling cannula was inserted into the sideport of an oxygen face-mask or was integrated as a separate line for nasal cannulas. The capnography waveform was displayed on the main physiological monitoring screen. A default ‘No breaths detected’ alert was triggered for apnea but no other audible or visual alarms were set for the capnography monitor. No restrictions or specific instructions were provided to clinicians regarding the detection of capnography waveform abnormalities as part of the research protocol because the study used observational design.

### Data collection

Demographic data and clinical characteristics were collected from medical records and self-report questionnaires prior to procedures. Intra-procedural data were collected in real-time by the researcher who was present in the procedure room. Direct observation of the participant was required to record the timing of sedation administrations and any interventions applied by sedation providers.

### Predictor variables

### Statistical analysis

The best model was chosen with respect to the optimal balance between maximising the ability to predict prolonged events and alarm frequency. The alarm frequency metric used was the ratio of correct predictions of short apneic periods to the total number of apneic periods that triggered alerts, which was calculated as TP/(TP + FP + apneas > 30seconds).

#### Modelling

#### Decision curve analysis

We used the net benefit decision analytic measure to assist with deciding whether using the models in practice would lead to better outcomes on average than the default capnography alarm management strategies currently used in practice. The default strategies are: 1) the aggressive approach, which involves triggering an alarm after brief periods of apnea (typically 15 seconds); and 2) the concervative approach, which involves triggering an alarm for only prolonged periods of apnea (typically 30 seconds). The net benefit is calculated using the formula:

$$
NB = \frac{TP}{N} - \frac{FP}{N} \times \frac{Pt}{1-Pt}
$$

Where $NB$ is the net benefit, $TP$ is the total number of true positives (apneic event at 15 seconds which was predicted to be prolonged and did persist for >30 seconds), $FP$ is the total of number of false positives (apneic events at 15 seconds that was predicted to be prolonged but did not persist for >30 seconds), $N$ is the sample size and $Pt$ is the probability threshold.

This formula essentially transforms the total number of true positives and false positives into a standardized scale, weighted by the relative harm or a false-positive result.[@vickers2006decision] For example, a net benefit of 0.07 means the net benefit of using the model would be 7 true positives from every 100 predictions from the model. This net benefit can result from any combination of true positives and false positives.[@vickers2016net] The higher the value of $Pt$, the more importance is placed on the number of false positives.

A decision curve was created to plot net benefits across a range of probability thresholds. The decision curve was compared against the aggressive alarm management strategy (alarm triggered at 15 seconds of apnea) and the conservative alarm management strategy (alarm triggered at 30 seconds of apnea). The decision curve takes into account a plausbile range of clinician preferences for the point at which an intervention should be initiated to treat apnea (this could be, for example, something as simple as verbal stimulation). We tested thresholds in the range from 0.4 to 0.8. In a practical sense, this means that we decided that all clinicians would not ever accept a probability for prolonged apnea lower than 0.4 as a useful trigger for intervention, and, conversely, that all clinicians would always consider that intervention is indicated if the probability of prolonged apnea was higher than 0.8. These probability thresholds can be interpreted as value preferences in the context of decision curve analysis. For example, a clinician who is 'risk-averse' may elect for a more conservative probability threshold (closer to 0.4). Individual participant characteristics will also influence clinicians' decisions about probability thresholds. A clinician may elect to intervene when the probability for prolonged apnea is 0.5 for an elderley patient with multiple comorbidities but not for a young patient who may be more likely to be able to tolerate longer periods of apnea.

A threshold of 0.5 inherently indicates that avoiding a false positive is as valuable of an outcome as identifiying a true positive. Preferences for probability thresholds below 0.5 are weighted such that identifying a true positive is more valuable than avoiding a false positive. Preferences for probability thresholds above 0.5 are weighted such that avoiding a false positive is more valuable than identifying a true positive.  Decision curves can be interpreted such that the strategy with the highest net benefit at each probability threshold has the highest clinical value.[@vickers2016net]

## Results

### Model developmenent and evaluation

Out of sample prediction accuracy for the models using cross-validation is summarized in Table 1. The best model, in terms of

### Decision curve analysis

Net benefit associated with the random forest model exceeded the strategy of intervening for all patients (the default strategy for the 'aggressive alarm management approach') across all probability thresholds from the range of 0.4 to 0.8 (Figure 2A). The interpretation is that the best clinical outcome would be achieved for clinicians who are willing to initiate an intervention to treat apnea at the 15 second mark if the probability of it being prolonged was more than 40%, by using the random forest model. The net benefit associated with the random forest model was higher than the strategy of not intervening for any patients at the 15 second mark (the default strategy for the 'conservative alarm management approach') across all probability thresholds from the range of 0.4 to 0.8 (Figure 2B).

## Discussion

Results from prior research indicated that using information about the history of previous respiratory states may be a promising solution in regards to predicting durations of apneic periods. It was found in a study of capnography waveform abnormalities during nurse-administered sedation that the hazard for apnea increased two-fold (HR 2.14, 95% CI 1.75 - 2.62) when a patient was in a state of hypoventilation (defined as >10% reduction in ETCO~2~ from baseline).[@conway2019pre] The risk of apnea also increased with each additional sedation dose (HR 2.86; 95% CI 2.15 - 3.81).[@conway2019pre] Results from an earlier study in a different population also supported the observations that the onset of apneic periods during sedation are associated with the history of previous respiratory states. Krauss and colleagues[@kraussCharacteristicsPredictorsApnea2016] used survival analysis to model the time to first apneic events in a sample of 312 patients undergoing procedural sedation with propofol or ketamine in the emergency department. It was identified that having an abnormal ETCO~2~ measurement 30, 60 and 90 seconds prior to an apneic event increased risk for apnea (hazard ratios with 95% CI for the different lag durations of 2.45 (1.63 - 3.69), 1.88 (1.21 - 2.92) and 2.06 (1.36 - 3.11), respectively).

We leveraged these preliminary findings about the associations between apneic events and the history of previous respiratory states in the present study by building a predictive model using a machine learning approach.

Using the machine learning model would produce a better clinical outcome, in terms of the balance between true positives and false positives for detection of apneic periods over 30 seconds, than a default strategy of intervening for everyone who has an apneic period ≥15 seconds.

## References